\chapter{Abstract}
\label{chap:abstract}

Drasil~\cite{Drasil2021} is a software suite for generating software, with
particular focus for generating \ACF{scs} following the requirements described
in an abstract \ACF{srs} template. The template breaks up scientific knowledge
into various categories, and the abstracted variant of the template makes it
digestible for Drasil. A series of \acsp{dsl} are used to ``fill in'' the
template, from which Drasil is able to interpret an instance of, and configure a
generation procedure to generate usable software. The depth of knowledge
contained in the template's theories is key to understanding to what end we can
interpret and use the instantiated theories. To strengthen this depth, we create
a system for extending Drasil's known theory kinds and clearly explaining which
ones are usable in any specific context. Similarly, each theory kind contains a
particular subset of mathematical language that is relevant to them, and we act
on this information to restrict usable expression terms to their related
contexts. To further enrich the admissibility of expressions, we also make one
of the most critical subsets, that for concrete theory transcription,
\textit{type-safe} by building a \textit{bidirectional} type-checker that
catches all existing typing issues. In doing so, we were able to uncover many
typing issues in both the case studies generating code and those not. Finally,
as Drasil relies on a plethora of different \textit{types} of knowledge, it
needs a place to store them. Thus, we create a system to store any instance of
any type of knowledge in Drasils memory bank of knowledge, and enforce a basic
set of requirements that each type of knowledge must satisfy.
