\begin{haskell}{ModelKinds}{curModelKinds}{https://github.com/JacquesCarette/Drasil/blob/dc3674274edb00b1ae0d63e04ba03729e1dbc6f9/code/drasil-theory/lib/Theory/Drasil/ModelKinds.hs\#L25-L47}
-- | Models can be of different kinds: 
--
--     * 'NewDEModel's represent differential equations as 'DifferentialModel's
--     * 'DEModel's represent differential equations as 'RelationConcept's
--     * 'EquationalConstraint's represent invariants that will hold in a system of equations.
--     * 'EquationalModel's represent quantities that are calculated via a single definition/'QDefinition'.
--     * 'EquationalRealm's represent MultiDefns; quantities that may be calculated using any one of many 'DefiningExpr's (e.g., 'x = A = ... = Z')
--     * 'FunctionalModel's represent quantity-resulting function definitions.
--     * 'OthModel's are placeholders for models. No new 'OthModel's should be created, they should be using one of the other kinds.
data ModelKinds e where
  NewDEModel            :: DifferentialModel -> ModelKinds e
  DEModel               :: RelationConcept   -> ModelKinds e -- TODO: Split into ModelKinds Expr and ModelKinds ModelExpr resulting variants. The Expr variant should carry enough information that it can be solved properly.
  EquationalConstraints :: ConstraintSet e   -> ModelKinds e
  EquationalModel       :: QDefinition e     -> ModelKinds e
  EquationalRealm       :: MultiDefn e       -> ModelKinds e
  OthModel              :: RelationConcept   -> ModelKinds e -- TODO: Remove (after having removed all instances of it).


-- | 'ModelKinds' carrier, used to carry commonly overwritten information from the IMs/TMs/GDs.
data ModelKind e = MK {
  _mk     :: ModelKinds e,
  _mkUID  :: UID,
  _mkTerm :: NP
}
\end{haskell}
