\prefacesection{Declaration of Academic Achievement}

This thesis was completed by Daniel M. Szymczak under the supervision of 
Jacques Carette and Spencer Smith. The principal original contributions of the 
author are the analysis, design, implementation, and evaluation presented here 
for the Drasil framework as a knowledge-centric approach to generating software 
artifacts from a centralized representation of domain knowledge. In particular, 
the author carried out the systematic analysis of redundancy and inconsistency 
across scientific software artifacts, developed the thesis account of Drasil's 
chunk- and recipe-based architecture, implemented substantial portions of the 
framework and its supporting representations, and reimplemented the case 
studies used in this thesis within Drasil to evaluate the approach.

This thesis also draws on prior and collaborative work. The broader Drasil 
project is part of an ongoing research effort, and the author's work should not 
be taken to mean that all ideas, infrastructure, or repository contents 
originated solely with the author. In addition, most of the case studies 
discussed in this thesis were originally developed by other teams at McMaster 
University; the author's contribution in this thesis is their analysis and 
re-expression within Drasil, rather than authorship of all original source 
artifacts for those systems. Although Drasil uses GOOL as an intermediate layer 
for code generation, GOOL itself was not developed by the author and is not 
claimed as a contribution of this thesis. Where prior templates, existing 
artifacts, supervisory guidance, or contributions from other project members 
informed the work, they are acknowledged as part of the broader research 
context so as to avoid misattribution.