\chapter{Future Work}
\label{c:future}
\ds{This can probably be a section at the end of results / conclusion instead?}

Development of Drasil is ongoing and the framework is still being iterated upon
to date. In this section we present areas we believe have room for improvement
along with plans for additional features to be added in the long-term.

\section{Inference}

\ds{something like the following}
- As mentioned earlier throughout Section~\ref{sec:iterefine}, and explicitly 
stated in Section~\ref{sec:glassBRSRSRecipe}, there are areas where Drasil can
stand to be refined further, particularly around inferring necessary chunks as 
opposed to explicitly defining them in lists. Taking the SRS as an example, we 
should be able to refine the system to the point where the System Information 
object can be simplified by inferring necessary chunks from the SRS Recipe 
definition. The current system is used namely due to the rigidity of Haskell's 
type system.
Figure~\ref{fig:codeSpec} - \codeH{constMap} is another example we should be 
able to infer eventually (hence the ``yet").

\section{Typed Expression Language}
- Will allow for much more in-depth sanity checking of generated \sfs{}

\section{Model types}

\section{Usability}
As mentioned in the results section, usability remains a great area for
improvement for Drasil. Work to create a visual front-end for the framework
has been planned and we hope to eventually get to the point of usability being
as simple as drag-and-drop or similar mechanisms.

Work on usability will address each of the core areas of development with
Drasil: knowledge-capture of system-agnostic information,
knowledge-capture of system-specific information, and recipe creation and
modification.

\ds{Developer experience plugins to improve usability would also be useful. 
Something like a VS Code extension}.

\section{Variants} 

\ds{may not need a section< I just don't want to forget this}

As we see in English, along with many other languages, there is no one way to 
define something. Having multiple variants of a natural language description 
for a chunk would be useful as each variant can be used in a different context 
for different effect to a given audience.

The Newton's Second Law example from Section~\ref{sec:chunky} shows a perfect 
example of this as the ``loosely translated" description is not the description 
we have chosen to encode in our chunk, yet both are equally valid.

\section{Many more artifacts/more document knowledge}

Journal papers, Jupyter notebooks, lesson plans, etc.

\section{More display variabilities}

\section{More languages (?)}

\ds{Programming and natural languages.}

\section{More scientific knowledge}

As Drasil sees more use, the knowledge-base will inevitably expand to encompass 
more knowledge within fields. We would also like to expand it into other 
scientific fields such as chemistry, medicine, etc.

\section{More computational knowledge}

Higher order ODEs, linear system solvers, external libs, etc.

More knowledge around software engineering and design processes.

\section{Measuring Drasil's Effectiveness}

We have made many observations as to how we believe Drasil can improve the 
lives of developers (Chapter~\ref{c:results}), however, we have not yet backed 
that up with hard data. We need to design several experiments with differing 
goals to test Drasil's effectiveness in improving developer quality of life. 
Here we propose several experiments:

\begin{itemize}
\item Test the difficulty of create net new software in Drasil vs traditional 
methods
\item Test the difficulty of finding and removing inconsistencies/errors using 
Drasil vs traditional methods
\item Comparing costs (incl. developer time) of maintenance using Drasil vs 
traditional methods
\end{itemize}

\ds{More TBD?}

\ds{Might want to rephrase the above into ``We want to create a research 
program to answer the following questions" and then list out the questions we 
want to answer, namely ``Is it easier to create software when using Drasil?", 
``Is it easier to find and correct errors when using Drasil?", and ``Are there 
significant time savings in long-term maintenance when using Drasil?"}
