With a focus on building Scientific Computing Software,
Drasil\thinspace\cite{Drasil2021} is an exploration of the ideology described in
\autoref{chap:ideology}. Rather than building one's software project in any
single or combination of general purpose programming language, the usage of a
sequence of domain-specific languages together in Drasil-based projects can be
used to describe common undergraduate level physics models and problems and
describe the target program that simulates said models.

\intodo{Last sentence of above: be more generic, and then mention that it is
      \textit{currently} focused on physics, etc.}

\section{An Exploration}

Originally known as Literate Scientific Software (LSS) \todo{cite}, Drasil is an
exploration of this ideology. Drasil's largest domain of knowledge covered
originates from LSS: scientific computing software (SCS). \porthref{Dr. Jacques
      Carette}{https://www.cas.mcmaster.ca/~carette/} and \porthref{Dr. Spencer
      Smith}{https://www.cas.mcmaster.ca/~smiths/} are the principal investigators of
Drasil. Deeply embedded in Haskell \todo{ref} (and currently built against
\acs{ghc} \todo{ref} 8.8.4), Drasil currently focuses on building tooling around
8 case studies, each following the formalized \ACF{srs} template
\cite{SmithAndLai2005}:

\intodo{Explain the focus of each case study in the list below:}
\begin{itemize}

      \item Single Pendulum

      \item Double Pendulum

      \item GamePhysics

      \item HGHC

      \item PDController

      \item SWHS

      \item NoPCM

      \item SSP

\end{itemize}

Drasil is currently capable of generating usable software through compiling to
\ACF{gool}, which is capable of producing Java, C++, Python, C\#
\cite{MacLachlan2020}, and Swift (not discussed in MacLachlan's Master's thesis,
but created by him as well, and available similarly). Drasil contains renderers
for HTML, Makefile, basic Markdown (enough for README), GraphViz DOT (graph
description language), plaintext documents, and \LaTeX\ / \TeX. Drasil's source
code is publicly available on
\porthref{GitHub}{https://github.com/JacquesCarette/Drasil}, and Drasil's
documentation
(\porthref{user-facing}{https://jacquescarette.github.io/Drasil/docs/index.html},
and
\porthref{internal}{https://jacquescarette.github.io/Drasil/docs/full/index.html})
is available on the Drasil project
\porthref{homepage}{https://jacquescarette.github.io/Drasil/}. A public Drasil
wiki is hosted on the \porthref{same GitHub
      project}{https://github.com/JacquesCarette/Drasil/wiki}, containing information
on potential future Drasil projects, Drasil-related papers, a
\porthref{developer workspace configuration and ``quick start''
      guide}{https://github.com/JacquesCarette/Drasil/wiki/New-Workspace-Setup}, and a
guide for \porthref{building your own project with
      Drasil}{https://github.com/JacquesCarette/Drasil/wiki/Creating-Your-Project-in-Drasil}.

\section{Methodology}

\begin{itemize}

      \item Primarily focused on undergraduate-level science (primarily physics)
            problems

      \item Generating scientific software artifacts
            \begin{itemize}
                  \item Develop a stable knowledge base for physics problems

                  \item Develop a stable framework for laying cookie-cutter
                        problems.

                  \item Make it as simple as using a projectional editor.

            \end{itemize}

      \item Approach to encoding knowledge -- e.g., ``bottom-up''

\end{itemize}

\section{State of Architecture}

\begin{itemize}

      \item Knowledge Encodings \& Organization
            \begin{itemize}
                  \item Chunks
                  \item ChunkDB
            \end{itemize}

      \item SmithEtAl template for SRS documents

      \item Haskell/GHC 8.8.4

      \item Uses a series of \ACFP{dsl} to build up the knowledge-base and
            ``story'' of a scene/project, until sufficient knowledge is built such
            that a generator may take the whole story and generate
            representational software artifacts.
            \intodo{Should add an example of the encoding of a theory.}

      \item Generates \acs{oo} programs; \acs{gool} is used to generate code in
            C++/C, Java, Python, Swift, and C\#.

      \item Doxygen, Markdown, HTML, TeX, Makefile, CSS, and more non-programming
            software artifacts.

      \item Creates a ``calculation path'' for a series of $x = f(a,b,c,...)$
            formed equations (e.g., simple LHS = complex RHS), and ODEs supported
            by python libraries.

      \item \intodo{table of Case Study/Generates SRS/Code/Complete \checkmark{}s to fill in the boxes}

      \item Multiple code generating examples (each with their own requirements):
            \begin{itemize}
                  \item GlassBR
                  \item NoPCM
                  \item PDController
                  \item Projectile
            \end{itemize}

      \item The calculation path relies (implicitly) on:
            \begin{itemize}

                  \item The Exprs of the RelationConcept being of the form: $x =
                              f(a,b,c,...)$

                  \item The ODEs being described by the expression language, but we
                        do not rely on being able to generate representing
                        computational code due to needing extra external information
                        (e.g., we require a supplementary packet of information
                              [ODEInfo]).

            \end{itemize}

\end{itemize}

\intodo{Short-term problems -- leading into topics}


