With a focus on building Scientific Computing Software, Drasil is an exploration
of this idea. Rather than building one's software project in any single or
combination of general purpose programming language, the usage of a sequence of
domain-specific languages together in Drasil-based projects can be used to
describe common undergraduate level physics models and problems and describe the
target program that simulates said models.

\begin{itemize}
    \item \porthref{public project
              webpage}{https://jacquescarette.github.io/Drasil/}
    \item \porthref{source code is hosted on
              GitHub}{https://github.com/JacquesCarette/Drasil}
    \item \porthref{public wiki is available on the same GitHub
              project}{https://github.com/JacquesCarette/Drasil/wiki}
          \begin{itemize}
              \item Specifically, \porthref{a workspace and ``quick start'' is
                        available on the
                        wiki.}{https://github.com/JacquesCarette/Drasil/wiki/New-Workspace-Setup}
          \end{itemize}
    \item Principal investigators: \porthref{Dr. Jacques
              Carette}{https://www.cas.mcmaster.ca/~carette/} \& \porthref{Dr.
              Spencer Smith}{https://www.cas.mcmaster.ca/~smiths/}
\end{itemize}

\section{Focus}

\begin{itemize}
    \item Primarily for undergraduate-level science (primarily physics) problems
    \item Generating scientific software artifacts
          \begin{itemize}
              \item Develop a stable knowledge base for physics problems
              \item Develop a stable framework for laying cookie-cutter
                    problems.
              \item Make it as simple as using a projectional editor.
          \end{itemize}
\end{itemize}

\section{Methodology}

\begin{itemize}
    \item Approach to encoding knowledge -- e.g., ``bottom-up''
\end{itemize}

\section{Architecture}

\begin{itemize}
    \item Knowledge Encodings
          \begin{itemize}
              \item Chunks
              \item ChunkDB
          \end{itemize}
    \item SmithEtAl template for SRS documents
    \item Haskell/GHC 8.8.4
\end{itemize}

\section{State}

\begin{itemize}
    \item Generates OO programs; GOOL is used to generate code in C++/C, Java,
          Python, Swift, and C\#.
          \intodo{non-code artifacts}
    \item Creates a ``calculation path'' for a series of $x = f(a,b,c,...)$
          formed equations (e.g., simple LHS = complex RHS), and ODEs supported
          by python libraries.
    \item Multiple examples:
          \begin{itemize}
              \item DblPendulum
              \item Game physics
              \item GlassBR
              \item HGHC
              \item NoPCM
              \item PDController
              \item Projectile
              \item SglPendulum
              \item SSP
              \item SWHS
          \end{itemize}
    \item Multiple code generating examples (each with their own requirements):
          \begin{itemize}
              \item GlassBR
              \item NoPCM
              \item PDController
              \item Projectile
          \end{itemize}
\end{itemize}

\intodo{Short-term problems -- leading into topics}
