\chapter{Introduction} \label{chap:intro}
Scientific computing (SC) is an intersection of computer science, mathematics, 
and science. It is a field that solves complex scientific problems by using 
computing techniques and tools. Writing documentation is a part of the process 
of developing scientific software. The role of documentation is to help 
people better understand the software and to ``communicate information to its 
audience and instil knowledge of the system it describes" 
\cite{forward2002software}. The significance of software documentation has 
been presented in many papers by previous researchers \cite{parnas2011precise}, 
\cite{chomal2014significance}, \cite{kipyegen2013importance}. It is further 
shown by Smith et al. \cite{SmithandKoothoor2016}, \cite{SmithandYu2007} that 
developing scientific computing software (SCS) in a document-driven methodology 
improves the quality of the software . 

Jupyter Notebook is a system for creating and sharing data science and 
scientific computing documentation. It is a nonprofit, open-source application 
born out in 2014, providing interactive computing across multiple programming 
languages, such as Python, Javascript, Matlab, and R. A Jupyter Notebook 
integrates text, live code, equations, computational outputs, visualizations, 
and multimedia resources, including images and videos. Jupyter Notebook is one 
of the most widely used interactive systems among scientists. Its popularity 
has grown from 200,000 to 2.5 million public Jupyter Notebooks on GitHub in 
three years from 2015 to 2018 \cite{Jeffrey2018}. It is used in a variety of 
areas and ways because of its flexibility and added values. For example, the 
notebook can be used as an educational tool in engineering courses, enhancing 
teaching and learning efficiency \cite{cardoso2019using}, \cite{zhao2019use}.

Even though the importance of documentation is widely recognized, it is often 
missing or poorly documented in SCS because: i) scientists are not aware of 
the why, how, and what of documentation \cite{hermann2022documenting}, 
\cite{chang2022understanding}; ii) it is time-consuming to produce 
\cite{sanders2008dealing}; iii) scientists generally believe that writing 
documentation demands more work and effort than they would likely yield in 
terms of the benefits of it \cite{smith2016advantages}.

We are trying to increase the efficiency of documentation development by 
adopting generative programming. Generative programming is a technique that 
allows programmers to write the code or document at a higher abstraction level, 
and the generator produces the desired outputs. Drasil is an application of 
generative programming, and it is the framework we use to conduct this 
research. Drasil saves us more time in the documentation development process by 
letting us encode each piece of information of our scientific problems once and 
generating the document automatically.


\section{Background}
\subsection{Drasil}
Drasil is a framework that can generate software artifacts, including Software 
Requirement Specifications (SRS), code (C++, C\#, Java, and Python), README, 
and Makefile, from a stable knowledge base. The goals of Drasil are reducing 
knowledge duplication and improving traceability \cite{drasil}. Drasil captures 
the knowledge through our hand-made case studies. We currently have 10 case 
studies that cover different physics problems, such as Projectile and Pendulum. 
Recipes for scientific problems are encoded in Drasil, and it generates code 
and documentation for us. Each piece of information only needs to be provided 
to Drasil once, and that information can be used wherever it is needed. By 
defining and storing common concepts in a central repository, and case-specific 
concepts in their own packages, Drasil enables the reuse of information across 
different engineering domains and applications. This feature significantly 
reduces the time and effort required for software development and 
documentation, while also improving the consistency and accuracy of the 
information being used. More information and an example of how knowledge is 
reused can be found in Chapter~\ref{chap:projMotion}. 

Drasil is currently capable of generating SRS, a template for designing and 
documenting scientific computing software requirement decisions created by 
Smith et al \cite{smith2005new}, in document languages HTML and LaTeX. We are 
looking to extend the capability of Drasil by generating Jupyter Notebook in 
Drasil.

\subsection{Jupyter Notebook}
Jupyter Notebook is an interactive open-source web application for creating and 
sharing computational science documentation that contains text, executable 
code, mathematical equations, graphics, and visualizations.

\subsubsection{Structure of a notebook document}
A Jupyter Notebook has two components: front-end ``cells" and back-end 
``kernels". The notebook consists of a sequence of cells: code cells, markdown 
cells, and raw cells. A cell is a multiline text input field. The notebook 
works by users entering a piece of information (text or programming code) in 
cells from the web page user interface. That information is then passed to the 
back-end kernels which execute the code and return the results 
\cite{notebookdoc}.

\subsubsection{The Value of Jupyter Notebook}
There are several advantages of Jupyter Notebook: sharable, all-in-one, and 
live code. First of all, the notebook is easy to share because it can be 
converted into other formats such as HTML, Markdown, and PDF. Secondly, it 
combines all aspects of data in one single document, making the document easy 
to visualize, maintain and modify. In addition, Jupyter Notebook provides an 
environment of live code and computational equations. Usually, when programmers 
are running code on some other IDEs, they have to write the entire program 
before executing it. However, the notebook allows programmers to execute a 
specific portion of the code without running the whole program. The ability to 
run a snippet of code and integrate with text highlight the usability of the 
notebook.

\section{Problem Statement}
Since both Jupyter Notebook and Drasil focus on creating and generating 
scientific computing documentation, we are interested in extending the values 
of Jupyter Notebook to Drasil and the kind of knowledge we can manipulate. 
Following are the three main problems we are trying to solve with Drasil in 
this paper:

\begin{enumerate}
	\item Generate Jupyter Notebooks. To acheive this, we will have 
	to generate documents in notebook format. Jupyter Notebook is a simple 
	JSON document with a .ipynb file extension. Notebook contents are either 
	code or Markdown. Therefore, non-code contents must be in Markdown format 
	with JSON layout. Drasil can only write in HTML and LaTeX. We are building 
	a notebook printer in Drasil for generating documents that are 
	readable and writable in Jupyter Notebook.
	\item Develop the structure of lesson plans and generate them. As 
	mentioned, Jupyter Notebook is used as an educational tool for teaching 
	engineering courses. When it comes to teaching, lesson plans are often 
	brought up because they help teachers to organize the daily activities 
	in each class time. We are interested in teaching Drasil a ``textbook" 
	structure by starting with generating a simple physics lesson plan and 
	expanding Drasil's application. We aim to capture the elements of 
	textbook chapters, identify the family of lesson plans, and classify the 
	knowledge to build a general structure in Drasil, which will enable the 
	lesson plan to generalize to a variety of lessons.
	\item Generate notebooks that mix text and code. Jupyter Notebook is 
	an interactive application for creating documents that contain formattable 
	text and executable code. However, Drasil doesn't support interactive 
	recipes. There is no code in SRS documents, and text and code are generated 
	separately in Drasil. We are looking for the possibility of generating a 
	notebook document that incorporate both text and code, thereby enhancing 
	the capabilities of Drasil and its potential to solve more scientific 
	problems.
\end{enumerate}

\section{Thesis Outline}
Thesis outline here.