\begin{longtable}[c]{|>{\raggedright}p{0.3\linewidth}|>{\raggedright\arraybackslash}p{0.54\linewidth}|}
    \caption{Drasil Case Studies}
    \label{tab:drasilCaseStudies}                                              \\

    \hline

    \rowcolor{McMasterMediumGrey}
    \textbf{Case Study}      & \textbf{Focus}
    \\ \hline

    \caseStudy{glassbr}      & {Predicting likelihood of a glass slab
            resisting a specified blast.}
    \\ \hline

    \caseStudy{projectile}   & {Determining if a launched projectile hits a
            target, assuming no flight collisions.}
    \\ \hline

    \caseStudy{sglpendulum}  & {Observing the motion of a single pendulum.}
    \\ \hline

    \caseStudy{dblpendulum}  & {Observing the motion of a double pendulum.}
    \\ \hline

    \caseStudy{gamephysics}  & {Modelling of an open source 2D rigid body
            physics library used for games.}
    \\ \hline

    \caseStudy{hghc}         & {Examining the heat transfer coefficients
            related to clad.}
    \\ \hline

    \caseStudy{pdcontroller} & {Examining the output of a ``Power Plant''
            (Process Variable) over time.}
    \\ \hline

    \caseStudy{swhs}         & {Modelling of a solar water heating system with
            phase change material, predicting temperatures and change in heat energy
            of water and the PCM over time.}
    \\ \hline

    \caseStudy{nopcm}        & {Modelling of a solar water heating system
            without phase change material, predicting temperatures and change in heat
            energy of water and the PCM over time.}
    \\ \hline

    \caseStudy{ssp}          & {Assessment of the safety of a slope (composed
            of rock and soil) subject to gravity, identifying the surface
            most likely to experience slip and an index of its relative
            stability (factor of safety).}
    \\ \hline
\end{longtable}
