\prefacesection{Abstract}

Scientific Computing (SC) involves analyzing and simulating complex scientific 
and engineering problems using computing techniques and tools. To improve the 
understandability, maintainability, and reproducibility of SC software,  
documentation should be an integral part of the development process. Jupyter 
Notebook is a popular tool for developing SC software documentation, and is 
also used to enhance teaching and learning efficiency in engineering education 
due to its flexibility and benefits. Despite the importance of documentation, 
it is often missing or poorly executed in SC software for several reasons, 
including time-consuming. 

Drasil is a framework that aims to improve the efficiency of documentation 
development. By encoding each piece of information of scientific problems once 
and generating the document automatically, Drasil saves time in the 
documentation development process. We are interested in generating Jupyter 
Notebooks in Drasil to expand its applications, including generating 
educational documents, and extend the values of Jupyter Notebooks to Drasil.

To achieve this, we implement a JSON printer capable of generating Drasil 
software artifacts, such as SRS, in notebook format. This enables us to 
generate Jupyter Notebooks in Drasil, and generate educational documents, 
starting with lesson plans. We develop the structure of our lesson plans and 
designed the language of lesson plans in Drasil. Additionally, Jupyter 
Notebooks seamlessly integrate different content types with code, making them 
ideal for data research. We discuss two approaches and the implementations to 
combine text and code in our Drasil-generated Jupyter Notebooks. 

This paper achieves our main objectives of generating Jupyter Notebooks that 
integrate both text and code, as well as generating lesson plans using Drasil.