\documentclass[12pt]{article}

\usepackage{bm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{colortbl}
\usepackage{xr}
\usepackage{hyperref}
\usepackage{longtable}
\usepackage{xfrac}
\usepackage{tabularx}
\usepackage{float}
\usepackage{siunitx}
\usepackage{booktabs}

\hypersetup{
    bookmarks=true,         % show bookmarks bar?
      colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=red,          % color of internal links (change box color with linkbordercolor)
    citecolor=green,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=cyan           % color of external links
}
\newcommand{\NN}[1]{{\color{red}#1}}

\newcommand{\colZwidth}{1.0\textwidth}
\newcommand{\blt}{- } %used for bullets in a list
\newcommand{\colAwidth}{0.2\textwidth}
\newcommand{\colBwidth}{0.73\textwidth}
\newcommand{\colCwidth}{0.1\textwidth}
\newcommand{\colDwidth}{0.05\textwidth}
\newcommand{\colEwidth}{0.8\textwidth}
\newcommand{\colFwidth}{0.17\textwidth}
\newcommand{\colGwidth}{0.5\textwidth}
\newcommand{\colHwidth}{0.28\textwidth}
\newcounter{defnum} %Definition Number
\newcommand{\dthedefnum}{GD\thedefnum}
\newcommand{\dref}[1]{GD\ref{#1}}
\newcounter{datadefnum} %Datadefinition Number
\newcommand{\ddthedatadefnum}{DD\thedatadefnum}
\newcommand{\ddref}[1]{DD\ref{#1}}
\newcounter{theorynum} %Theory Number
\newcommand{\tthetheorynum}{T\thetheorynum}
\newcommand{\tref}[1]{T\ref{#1}}
\newcounter{tablenum} %Table Number
\newcommand{\tbthetablenum}{T\thetablenum}
\newcommand{\tbref}[1]{TB\ref{#1}}
\newcounter{assumpnum} %Assumption Number
\newcommand{\atheassumpnum}{P\theassumpnum}
\newcommand{\aref}[1]{A\ref{#1}}
\newcounter{goalnum} %Goal Number
\newcommand{\gthegoalnum}{P\thegoalnum}
\newcommand{\gsref}[1]{GS\ref{#1}}
\newcounter{instnum} %Instance Number
\newcommand{\itheinstnum}{IM\theinstnum}
\newcommand{\iref}[1]{IM\ref{#1}}
\newcommand{\tclad}{T_\text{CL}}
\newcommand{\degree}{\ensuremath{^\circ}}
\newcommand{\progname}{SWHS}

%\oddsidemargin 0mm
%\evensidemargin 0mm
%\textwidth 160mm
%\textheight 200mm
\usepackage{fullpage}

\begin{document}

\title{Verification and Validation Report for Solar Water Heating Systems
  Incorporating Phase Change Material} 
\author{Maya Grab and Brooks MacLachlan}
\date{\today}
	
\maketitle

\tableofcontents

\section{Introduction}
This document is a report on the results of a testing suite for a Solar Water 
Heating System incorporating Phase Change Material simulator.
\\Detailed descriptions of the tests executed can be found in the Verification 
and Validation Plan document.

\section{System Tests}

\subsection{Faulty Input}
The following tests have failed by verification:

\begin{center}
	\begin{longtable}{ | r | p{3cm} | p{4cm} | p{4cm} | p{4cm} | p{4cm}|}
	\caption{Faulty Input Cases} \\ \hline \label{TblInputVar} 
	No. & Input & Expected Outcome & Expected MsgID & Actual MsgID \\ \hline
	40 & $T_{\text{init}} = 45$ &error: Tinit must be $< 0$ Tmelt & input:TinitTmelt & MATLAB:badsubscript \\ \hline
	\end{longtable}
\end{center}

\subsection{Closed Form Solution for Latent Heating}
The relative error between the closed form solution for water temperature and the simulator solution was 0.0046.

\subsection{Alternative Algorithm}
When comparing the output of a simulator using the ode23 solver to this simulator (using ode45), there was zero difference in the results in all cases.

\subsection{Comparison to Original Implementation}
All tests passed with a relative tolerance of 0.002.


\section{Unit Tests}

\subsection{Temperature Modules}
All tests passed within a relative tolerance of 1e-06.

\subsection{Energy Modules}
All tests passed within a relative tolerance of 1e-08.

\subsection{Event Modules} 
All tests passed within a relative tolerance of 1e-08.

\end{document}